Checking Big Suffix and LCP Arrays by Probabilistic Methods

### 2017-04-25 draft/draft.tex
2017/04/25

\author{
	Yi~Wu,
	Ge~Nong,
	Wai~Hong~Chan,
	Ling~Bo~Han% <-this % stops a space
	\IEEEcompsocitemizethanks{
		\IEEEcompsocthanksitem Y. Wu, G. Nong (corresponding author) and L.~B.~Han are with the Department of Computer Science, Sun Yat-sen University, Guangzhou 510275, China. E-mails: wu.yi.christian@gmail.com, issng@mail.sysu.edu.cn, hanlb@mail2.sysu.edu.cn.

@@

For full-text indexing of massive data, the suffix and LCP (longest common prefix) arrays have been recognized as fundamental data structures, and there are at least two needs in practice for checking their correctness, i.e. program debugging and verifying the arrays constructed by probabilistic algorithms. In this paper, we propose two probabilistic methods to check the suffix and LCP arrays of constant or integer alphabets in external memory by using a Karp-Rabin fingerprinting technique, where the checking result is wrong only with a negligible error probability. The first method checks the lexicographical order and the LCP-value of two suffixes by computing and comparing the fingerprints of their LCPs. This method is rather general in terms of that it can verify any full or sparse suffix/LCP array of any order. The second method is more space efficient, it first employs the fingerprinting technique to verify a subset of the given suffix and LCP arrays, from which then a copy of the suffix and LCP arrays is produced using the induced sorting principle and compared with the given arrays for verification, where the copy of the induced suffix and LCP arrays can be removed for constant alphabets.

@@

Suffix and longest common prefix (LCP) arrays play an important role in various string processing tasks, such as data compression, pattern matching and genome assembly. In many applications, these two data structures make up the core part of a powerful full-text index, called enhanced suffix array~\cite{Abouelhodaa2004}, which is more space efficient than a suffix tree and applicable to emulating most searching functionalities provided by the latter in the same time complexity. The first algorithm for building suffix array~(SA) in internal memory was presented in~\cite{Manber1993}. From then on, much more effort has been put on designing efficient constructors for suffix array on different computation models~\cite{Karkkainen2003, Ko2003, Kim2003, Nong11, Dementiev2008, Ferragina2012, Manzini2004, Bingmann12, Karkkainen2014, Nong14, Nong15}. In respect of the research on LCP array construction algorithms, the existing works can be classified into two categories with regard to their input requirements, where the algorithms from the first category compute both suffix and LCP arrays at the same time with the original text only~\cite{Fischer11, Bingmann12, Flick2015}, and those from the second category carry out the computation by taking SA and/or Burrows-Wheeler transform (BWT) as additional inputs~\cite{Kasai2001,Karkkainen2009, Fischer11, Puglisi2008, Deo2013}. So far, the algorithms designed by the induced sorting~(IS) principle take linear time and space to run and get the best results on both internal and external memory~\cite{Nong11, Karkkainen2014, Karkkainen2017}. In addition to the sequential algorithms, there are also parallel algorithms proposed to achieve high performance by fully using the available multi-core CPUs and/or GPUs~\cite{Osipov2012, Deo2013, Wang2015, Karkkainen2015, Karkkainen2016}.

@@

Our contribution comprises two methods to probabilistically verify any given suffix and LCP arrays. In principle, Method A checks the lexical order and the LCP-value of two neighboring suffixes in the suffix array by literally comparing the characters of their LCPs. For reducing the time complexity of a comparison between two sequences of characters, we use a Karp-Rabin fingerprinting technique to convert each sequence into a single integer, called fingerprint, and compare the fingerprints instead to check the equality of two sequences. The algorithm for Method A involves multiple scans and sorts on sets of $\mathcal{O}(n)$ fixed-size items. Its implementation in external memory suffers from a space bottleneck due to the large disk volume taken by each sort. 
To overcome this drawback, Method B first employs the fingerprinting technique to check a subset selected from the given suffix and LCP arrays, then it reuses the inducing process of the IS-based construction algorithm SA-IS \cite{Nong11} to produce the final suffix and LCP arrays from the verified subset and literally compares them with the input arrays to ensure the correctness of the latter. Our experiments indicate that the peak disk use of the programs for Algorithm~\ref{alg:2} designed by Method B is only about half as that of the program for Algorithm~\ref{alg:1} designed by Method A.

The remainder of this paper is organized as follows. Sections~\ref{sec:method1} and~\ref{sec:method2} describe the two methods and their algorithmic designs. Section~\ref{sec:experiment} conducts an experimental study for performance evaluation of our programs for these algorithms. Section~\ref{sec:conclusion} gives some concluding remarks.

@@

\begin{IEEEproof}
	Both the sufficiency and necessity are immediately seen from the definition of suffix and LCP arrays. Specifically, condition (1) demonstrates that all the suffixes in $x$ are sorted in $sa$, while conditions (2)-(3) indicate that the lexical order and the LCP-value of any two neighboring suffixes in $sa$ are both correct.
\end{IEEEproof}

If we directly compare the characters of two suffixes, the worst case time is $O(n)$. An alternative is to exploit a perfect hash function to convert each substring into a single integer such that any two substrings have a common hash value if and only if they are literally equal to each other, hence we can compare the hash values of two substrings instead to check their equality. Taking into account the high difficulty of finding such a perfect hash function to meet this requirement, we prefer using a Karp-Rabin fingerprinting function~\cite{Karp1987} to transform a substring into an integer called fingerprint. To be specific, suppose $L$ is a prime and $\delta$ is a number randomly chosen from $[1, L)$, the fingerprint ${\sf fp}(i, j)$ for a substring $x[i, j]$ can be iteratively calculated according to the formulas below as follows: scan $x$ rightward to iteratively compute ${\sf fp}(0, k)$ for all $k \in [0, n)$ using Formulas~\ref{formula:1}-\ref{formula:2}, record ${\sf fp}(0, i - 1)$ and ${\sf fp}(0, j)$ during the calculation and subtract the former from the latter to obtain ${\sf fp}(i, j)$ using Formula~\ref{formula:3}.

@@

Notice that two equal substrings always share a common fingerprint, but the inverse is not true. It has been proved in~\cite{Karp1987} that the probability of a false match can be reduced to a negligible level by setting $L$ to a large value\footnote{This property is utilized in~\cite{Bille2013} to design a probabilistic algorithm for computing a sparse suffix array.}. Hence, we have:

\begin{Corollary} \label{corollary:1}
	Both $sa[0, n)$ and $lcp[0, n)$ are correct with a high probability given the following conditions, for all $i \in [1, n)$:

@@

Fig.~\ref{fig:example} gives an illustrating example for utilizing Corollary~\ref{corollary:1} to check the input suffix and LCP arrays. Given that $L = 197$ and $\delta = 101$, lines 4-8 compute ${\sf fp}(0, p)$ iteratively according to Formulas~\ref{formula:1}-\ref{formula:2}. Then, lines 10-16 use these values to compute the fingerprints for all the target substrings. In more detail, consider the leftmost pair of neighboring suffixes in $sa$, that is ${\sf suf}(sa[0])$ and ${\sf suf}(sa[1])$, the substrings indicated by their LCP-value are $x[sa[0], sa[0] + lcp[1] - 1]$ and $x[sa[1], sa[1] + lcp[1] - 1]$, respectively. According to Formula~\ref{formula:3}, ${\sf fp}(sa[0], sa[0] + lcp[1] - 1)$ is equal to the difference between ${\sf fp}(0, sa[0] - 1)$ and ${\sf fp}(0, sa[0] + lcp[1] - 1)$, both of which have been calculated beforehand. Following the same way, ${\sf fp}(sa[1], sa[1] + lcp[1] - 1)$ is computed by reducing ${\sf fp}(0, sa[1] - 1)$ from ${\sf fp}(0, sa[1] + lcp[1] - 1)$. Hence, we obtain the fingerprints for these two substrings in lines 10-12 and see that they are equal to each other.

@@

	\includegraphics[width = 0.9\columnwidth]{example}
	\caption{An example for computing and comparing fingerprints for substrings specifed by the LCP-Values of neighboring suffixes in the suffix array. \label{fig:example}}	
\end{figure}

@@

It is clear that the above algorithm consumes $\mathcal{O}(n)$ time and space when implemented in internal memory. However, if the two auxiliary arrays cannot be wholly accommodated into RAM during the execution of S2, it suffers from a performance degradation caused by frequent random accesses to disks. Assume that $x$, $sa$ and $lcp$ are stored in external memory, we design Algorithm~\ref{alg:1} for conducting these I/O operations in a disk-friendly way. The main idea is to first sort data in their access order and then visit them by sequential reads. For the purpose, Algorithm~\ref{alg:1} first scans $sa$ and $lcp$ to produce $ST_1, ST_2, ST_3$ and sorts their tuples by 1st component in ascending order at the very beginning (lines~\ref{alg:1:a}-\ref{alg:1:b}). Afterward, it iteratively computes the fingerprints of all the prefixes according to Formulas~\ref{formula:1}-\ref{formula:2} and assigns them to the sorted tuples as following (lines~\ref{alg:1:c}-\ref{alg:1:d}): when figuring out ${\sf fp}(0, i - 1)$, extract each tuple $e$ with $e.1st = i$ from $ST_1/ST_2/ST_3$, update $e$  with ${\sf fp}(0, i - 1)$, and then forward $e$ to $ST_1'/ST_2'/ST_3'$. Because the 1st components of the tuples in $ST_1$ constitute a copy of $sa$, the algorithm checks condition (1) when scanning these tuples in their sorted order. Finally, it sorts the updated tuples back to their original order~(line~\ref{alg:1:e}) and visits them  sequentially to check conditions (2)-(3) following the same way of S2~(lines~\ref{alg:1:f}-\ref{alg:1:g}).

@@

		sort tuples in $ST_1$, $ST_2$ and $ST_3$ by 1st component \label{alg:1:b}

@@

			\If{$ST_1.{\sf notEmpty}()$ {\sf and} $ST_1.{\sf top}().1st = i$}{~\label{alg:1:h}
				$e := ST_1.{\sf top}()$, $ST_1.{\sf pop}()$, $e.3rd := fp$, $ST_1'.{\sf push}(e)$
			}
			\Else{
				\Return false \hspace{5cm} // condition (1) is violated
			}~\label{alg:1:i}
			\While{$ST_2.{\sf notEmpty}()$ {\sf and} $ST_2.{\sf top}().1st = i$}{

@@

			$fp := fp \cdot \delta + x[i] \! \mod \! P$ \hspace{5cm} // $x[n]$ is the virtual character
		} \label{alg:1:d}

		sort tuples in $ST_1'$, $ST_2'$ and $ST_3'$ by 2nd component.  \label{alg:1:e}

@@

				\Return false \hspace{5cm} // condition (2) or (3) is violated
			}	
		} \label{alg:1:g}

@@

The last point to be mentioned here is how to obtain the value of $\delta^{lcp[i]}$ quickly when computing $\hat{fp_1}$ and $\hat{fp_2}$ in lines~\ref{alg:1:j} and~\ref{alg:1:k}. One method is to keep a lookup table in internal memory to store $\delta^{lcp[i]}$ for all $i \in [0, n)$. This can answer the question in constant time, but it is impractical if the table is larger than the available memory size. Notice that the LCP of any two suffixes is smaller than $n$, thus we can return the answer in $\mathcal{O}(\lceil \log_2{n} \rceil)$ time using $\mathcal{O}(\lceil \log_2{n} \rceil)$ internal memory space based on the following idea. Suppose $e$ is an integer from $[0, n)$, it can be broken down into the form of $\Sigma_{i = 0}^{\lceil \log_2{n} \rceil}{k_i \cdot 2^i}$ by performing at most $\lceil \log_2{n} \rceil$ divisions, where $k_0, k_1, ..., k_{\lceil \log_2{n} \rceil}$ are integers from $\{0, 1\}$ . Thereby, we have $\delta^e = \Pi_{i = 0}^{\lceil \log_2{n} \rceil}{\delta}^{k_i \cdot 2^i}$ by replacing $e$ with its decomposition, where the expression on the right side of the equation can be easily computed with $\{{\delta}^{1}, {\delta}^{2}, \dots, {\delta}^{2^{\lceil \log_2{n} \rceil}} \}$ already known.

@@

Algorithm~\ref{alg:1} performs multiple scans and sorts on the arrays of $\mathcal{O}(n)$ fixed-size tuples residing on disks. Given an external memory model with RAM size $M$, disk size $D$ and block size $B$, all are in words, the time and I/O complexities for each scan are $\mathcal{O}(n)$ and $\mathcal{O}(n / B)$, respectively, while those for each sort are $\mathcal{O}(n\log_{M/ B}(n / B))$ and $\mathcal{O}((n / B)\log_{M / B}(n / B))$, respectively~\cite{Arge2013}. Algorithm~\ref{alg:1} reaches its peak disk use when sorting tuples in lines~\ref{alg:1:b} and~\ref{alg:1:e}. Suppose the input string and the suffix/LCP array are encoded using $\alpha$- and $\beta$-byte integers, respectively, and each fingerprint is represented by a $\gamma$-byte integer, it takes $(2 \cdot (2 \cdot \alpha + \beta)) \cdot n$ space for sorting $ST_1/ST_1'$ and $(2 \cdot (2 \cdot \alpha + \beta + \gamma)) \cdot n$ for sorting $ST_2/ST_2'$ and $ST_3/ST_3'$. To save disk space, our program implementing the algorithm sorts $ST_1/ST_1'$, $ST_2/ST_2'$ and $ST_3/ST_3'$ separately and performs a single scan over $x$ for each of them to obtain the target fingerprints, resulting in smaller space requirement but larger running time. The experimental study in Section~\ref{sec:experiment} indicates that this program is still space consuming, its peak disk use is 40 bytes per input character.

@@

In this part, we describe an alternative checking method based on the induced sorting principle. Compared with Algorithm~\ref{alg:1}, our programs for Algorithm~\ref{alg:2} designed by this method take around half disk space on real-world datasets. Before the presentation, we introduce some symbols and notations used in the following paragraphs.

{\em Character and suffix classification.} All the characters in $x$ are classified into three types, namely L-, S- and S*-type. In detail, $x[i]$ is L-type if (1) $i = n - 1$ or (2) $x[i] > x[i + 1]$ or (3) $x[i] = x[i + 1]$ and $x[i + 1]$ is L-type; otherwise, $x[i]$ is S-type. Further, if $x[i]$ and $x[i + 1]$ are separately L-type and S-type, then $x[i + 1]$ is also an S*-type character. Moreover, a suffix is L-, S- or S*-type if its heading character is L-, S-, or S*-type, respectively.

{\em Suffix and LCP buckets.} Suppose $sa$ is correct, all the suffixes in $sa$ are naturally partitioned into multiple buckets and those of a common heading character are grouped into a single bucket that occupies a contiguous interval in $sa$. Each bucket can be further divided into two sub-buckets, where the left and the right parts only contain L- and S-type suffixes, respectively. For short, we use ${\sf sa\_bkt}(c)$ to denote the bucket storing suffixes starting with character $c$ and ${\sf sa\_bkt_L}(c)/{\sf sa\_bkt_S}(c)$ to denote its left/right sub-bucket. Accordingly, $lcp$ can be also split into multiple buckets, where ${\sf lcp\_bkt}(c)/{\sf lcp\_bkt_L}(c)/{\sf lcp\_bkt_S}(c)$ stores the LCP-values of suffixes in ${\sf sa\_bkt}(c)/{\sf sa\_bkt_L}(c)/{\sf sa\_bkt_S}(c)$.

{\em Suffix and LCP arrays for S*-type suffixes.} Given that the number of S*-type suffixes is $n_1$, $sa^*[0, n_1)$ stores all the S*-type suffixes arranged in lexical order, while $lcp^*[0] = 0$ and $lcp^*[i]$ records the LCP-value of ${\sf suf}(sa^*[i])$ and ${\sf suf}(sa^*[i - 1])$ for $i \in [1, n_1)$.

{\em Type Array.} The array $t$ records in $t[i]$ the type information of $x[i]$, where $t[i] = 1$ or $0$ if $x[i]$ is S-type or L-type, respectively.

@@

The induced sorting principle has been extensively used to design efficient algorithms for constructing the suffix and LCP arrays on internal and external memory models. An IS-based construction algorithm mainly consists of a reduction phase for computing $sa^*$ and $lcp^*$, followed by an induction phase for inducing $sa$ and $lcp$ from $sa^*$ and $lcp^*$~\footnote{In the interest of completeness, we give an overview of the induction phase in Appendix~\ref{sec:appendix}.}. Given that $sa^*$ and $lcp^*$ are already known, we can produce the final suffix and LCP arrays by calling the inducing process of an existing IS-based construction algorithm. This enlightens us to design a checker based on Lemma~\ref{lemma:2}.

@@

\begin{enumerate}[(1)]
	\item
	$sa^*$ and $lcp^*$ are both correct.
	\item
	$sa = sa'$ and $lcp = lcp'$, where $sa'$ and $lcp'$ are induced from $sa^*$ and $lcp^*$ by the IS method.
\end{enumerate}
\end{Lemma}

Similar to Method A, the fingerprinting technique can be employed to probabilistically check the correctness of $sa^*$ and $lcp^*$. An external-memory algorithm for checking the conditions in Corollary~\ref{corollary:2} is given in the next subsection.

@@

	sort tuples in $ST_1$ by 1st component \label{alg:2:b}

@@

	sort tuples in $ST_2$ by the 2nd component \label{alg:2:g}

@@

{\bf Assume that the alphabet $\Sigma$ is of a constant size, we can check $sa$ and $lcp$ without producing a copy of the final suffix and LCP arrays in Algorithm~\ref{alg:2}.} The idea is to compare the induced suffix/LCP items with their corresponding items in $sa/lcp$ during the inducing process. Specifically, when a suffix/LCP item $v_1$ is induced into a bucket, we check if it is equal to the corresponding item $v_2$ in $sa/lcp$. If $v_1 = v_2$, then $v_2$ is correct and we further use this value to induce the remaining suffix/LCP items. The key point here is to retrieve items of $sa/lcp$ from external storage as fast as possible. This can be done by conducting sequential I/O operations if we provide a read pointer together with a buffer for each suffix/LCP sub-bucket. We describe below more details of the adapted inducing process, where $lp_1/lp_2$ and $sp_1/sp_2$ are file pointer arrays for reading items from the sub-buckets of $sa$ and $lcp$.

>>>>

\author{
	Yi~Wu,
	Ge~Nong,
	Wai~Hong~Chan,
	Ling~Bo~Han% <-this % stops a space
	\IEEEcompsocitemizethanks{
		\IEEEcompsocthanksitem Y. Wu, G. Nong (corresponding author) and L.~B.~Han are with the Department of Computer Science, Sun Yat-sen University, Guangzhou, and the SYSU-CMU Shunde International Joint Research Institute, Shunde, China. E-mails: wu.yi.christian@gmail.com, issng@mail.sysu.edu.cn, hanlb@mail2.sysu.edu.cn.

@@

For full-text indexing of massive data, the suffix and LCP (longest common prefix) arrays have been recognized as fundamental data structures, and there are at least two needs in practice for checking their correctness, i.e. program debugging and verifying the arrays constructed by probabilistic algorithms. Two probabilistic methods are proposed to check the suffix and LCP arrays of constant or integer alphabets in external memory using a Karp-Rabin fingerprinting technique, where the checking  is wrong only with a negligible error probability. The first method checks the lexicographical order and the LCP-value of two suffixes by computing and comparing the fingerprints of their LCPs. This method is general in terms of that it can verify any full or sparse suffix/LCP array of any order. The second method uses less space, it first employs the fingerprinting technique to verify a subset of the given suffix and LCP arrays, from which two new suffix and LCP arrays are induced and compared with the given arrays for verification, where the induced suffix and LCP arrays can be removed for constant alphabets to save space.

@@

Suffix and longest common prefix (LCP) arrays play an important role in various string processing tasks, such as data compression, pattern matching and genome assembly. In many applications, these two data structures make up the core part of a powerful full-text index, called enhanced suffix array~\cite{Abouelhodaa2004}, which is more space efficient than a suffix tree and applicable to emulating most searching functionalities provided by the latter in the same time complexity. The first algorithm for building suffix array~(SA) in internal memory was presented in~\cite{Manber1993}. From then on, much more effort has been put on designing efficient constructors for suffix array on different computation models~\cite{Karkkainen2003, Ko2003, Kim2003, Nong11, Dementiev2008, Ferragina2012, Manzini2004, Bingmann12, Karkkainen2014, Nong14, Nong15}. In respect of the research on LCP array construction algorithms, the existing works can be classified into two categories with regard to their input requirements, where the algorithms from the first category compute both suffix and LCP arrays at the same time with the original text only~\cite{Fischer11, Bingmann12, Flick2015}, and those from the second category carry out the computation by taking SA and/or Burrows-Wheeler transform (BWT) as additional inputs~\cite{Kasai2001,Karkkainen2009, Fischer11, Puglisi2008, Deo2013}. So far, the algorithms designed by the induced sorting~(IS) principle take linear time and space to run and get the best results on both internal and external memory~\cite{Nong11, Bingmann12}. In addition to the sequential algorithms, there are also parallel algorithms proposed to achieve high performance by fully using the available multi-core CPUs and/or GPUs~\cite{Osipov2012, Deo2013, Wang2015, Karkkainen2015, Karkkainen2016}.

@@

Our contribution comprises two methods to probabilistically verify any given suffix and LCP arrays. In principle, Method A checks the lexical order and the LCP-value of two neighboring suffixes in the suffix array by literally comparing the characters of their LCPs. For reducing the time complexity of a comparison between two sequences of characters, we use a Karp-Rabin fingerprinting technique to convert each sequence into a single integer, called fingerprint, and compare the fingerprints instead to check the equality of two sequences. The algorithm for Method A involves multiple scans and sorts on sets of $\mathcal{O}(n)$ fixed-size items. Its implementation in external memory suffers from a space bottleneck due to the large disk volume taken by each sort. 
To overcome this drawback, Method B first employs the fingerprinting technique to check a subset selected from the given suffix and LCP arrays, 
%then it reuses the inducing process of the IS-based construction algorithm SA-IS \cite{Nong11} to produce the final suffix and LCP arrays from the verified subset and literally compares them with the input arrays to ensure the correctness of the latter. Our experiments indicate that the peak disk use of the programs for Algorithm~\ref{alg:2} designed by Method B is only about half as that of the program for Algorithm~\ref{alg:1} designed by Method A. 
then it utilizes the IS method to produce the final suffix and LCP arrays from the verified subset and literally compares them with the input arrays to ensure the correctness of the latter. Our experiments indicate that the program for Algorithm~\ref{alg:2} designed by Method B only takes around half as much disk space as the program for Algorithm~\ref{alg:1} designed by Method A.

The remainder of this paper is organized as follows. Sections~\ref{sec:method1} and~\ref{sec:method2} describe the two methods and their algorithmic designs. Section~\ref{sec:experiment} conducts an experimental study for performance evaluation of our programs for these algorithms. Finally, Section~\ref{sec:conclusion} gives some concluding remarks.

@@

\begin{IEEEproof}
	Both the sufficiency and necessity are immediately from the definitions of suffix and LCP arrays. Condition 1 guarantees that each suffix is in $sa$, conditions 2 and 3 guarantees that the lexical order and the LCP-value of any two neighboring suffixes in $sa$ are both correct.
\end{IEEEproof}

Directly comparing the characters of two suffixes to determine their LCP has the worst case time of $O(n)$. An alternative is to exploit a perfect hash function to convert each substring into a single integer such that any two substrings have a common hash value if and only if they are literally equal to each other, hence the hash values of two substrings can be compared instead to check the equality of two substrings. Taking into account the high difficulty of finding a perfect hash function to meet this requirement, we prefer using a Karp-Rabin fingerprinting function~\cite{Karp1987} to transform a substring into an integer called fingerprint. To be specific, suppose $L$ is a prime and $\delta$ is a number randomly chosen from $[1, L)$, the fingerprint ${\sf fp}(i, j)$ for a substring $x[i, j]$ can be iteratively calculated according to the formulas below: scan $x$ rightward to iteratively compute ${\sf fp}(0, k)$ for all $k \in [0, n)$ using Formulas~\ref{formula:1}-\ref{formula:2}, record ${\sf fp}(0, i - 1)$ and ${\sf fp}(0, j)$ during the calculation and subtract the former from the latter to obtain ${\sf fp}(i, j)$ using Formula~\ref{formula:3}.

@@

Notice that two equal substrings always share a common fingerprint, but the inverse is not true. The probability of a false match can be reduced to a negligible level by setting $L$ to a large value~\cite{Karp1987}, this property is utilized in~\cite{Bille2013} to design a probabilistic algorithm for computing a sparse suffix array. Hence we have:

\begin{Corollary} \label{corollary:1}
	Both $sa[0, n)$ and $lcp[0, n)$ are correct with a high probability given these conditions, for all $i \in [1, n)$:

@@

Fig.~\ref{fig:example} gives an illustrating example for utilizing Corollary~\ref{corollary:1} to check the input suffix and LCP arrays. Given that $L = 197$ and $\delta = 101$, lines 4-8 compute ${\sf fp}(0, p)$ iteratively according to Formulas~\ref{formula:1}-\ref{formula:2}. Lines 10-20 use these values to compute the fingerprints for all the target substrings. In more detail, consider the leftmost pair of neighboring suffixes in $sa$, i.e. ${\sf suf}(sa[0])$ and ${\sf suf}(sa[1])$, the substrings given by their LCP-value are $x[sa[0], sa[0] + lcp[1] - 1]$ and $x[sa[1], sa[1] + lcp[1] - 1]$, respectively. According to Formula~\ref{formula:3}, ${\sf fp}(sa[0], sa[0] + lcp[1] - 1)$ is equal to the difference between ${\sf fp}(0, sa[0] - 1)$ and ${\sf fp}(0, sa[0] + lcp[1] - 1)$, both have been calculated. Following the same way, ${\sf fp}(sa[1], sa[1] + lcp[1] - 1)$ is computed by reducing ${\sf fp}(0, sa[1] - 1)$ from ${\sf fp}(0, sa[1] + lcp[1] - 1)$. Hence, we obtain the fingerprints for these two substrings in lines 10-14 and see that they are equal to each other.

@@

	\includegraphics[width = 0.9\columnwidth]{example}
	\caption{An example for computing and comparing fingerprints for substrings specified by the LCP-values of neighboring suffixes in $sa$. \label{fig:example}}	
\end{figure}

@@

The above approach takes $\mathcal{O}(n)$ time and space in internal memory. The difficulty for applying it in external memory is step S2, it suffers from a performance degradation caused by frequent random accesses to disks. Assume that $x$, $sa$ and $lcp$ are stored in external memory, we design Algorithm~\ref{alg:1} for conducting these I/O operations in a disk-friendly way. The idea is to first sort data in the order that they are to be visited and then access them sequentially. For this purpose, Algorithm~\ref{alg:1} first scans $sa$ and $lcp$ to produce $ST_1, ST_2, ST_3$, and sorts their tuples by the first components in ascending order at the beginning (lines~\ref{alg:1:a}-\ref{alg:1:b}). Afterward, it iteratively computes the fingerprints of all the prefixes of $x$ according to Formulas~\ref{formula:1}-\ref{formula:2} and assigns them to the sorted tuples as follows (lines~\ref{alg:1:c}-\ref{alg:1:d}): when figuring out ${\sf fp}(0, i - 1)$, extract each tuple $e$ with $e.1st = i$ from $ST_1/ST_2/ST_3$, update $e$  with ${\sf fp}(0, i - 1)$, and then forward $e$ to $ST_1'/ST_2'/ST_3'$. Because the first components of the tuples in $ST_1$ constitute a copy of $sa$, the algorithm checks condition 1 when scanning these tuples in their sorted order. Finally, it sorts the updated tuples back to their original order~(line~\ref{alg:1:e}) and visits them sequentially to check conditions 2 and 3 following the same way of step S2~(lines~\ref{alg:1:f}-\ref{alg:1:g}).

@@

		sort the tuples in $ST_1$, $ST_2$ and $ST_3$ by the 1st components, respectively; \label{alg:1:b}

@@

			\If{$ST_1.{\sf notEmpty}()$ {\sf and} $ST_1.{\sf top}().1st = i$}{~\label{alg:1:h}
				$e := ST_1.{\sf top}()$, $ST_1.{\sf pop}()$, $e.3rd := fp$, $ST_1'.{\sf push}(e)$
			}
			\Else{
				\Return false \hspace{1cm} // condition 1 is violated
			}~\label{alg:1:i}
			\While{$ST_2.{\sf notEmpty}()$ {\sf and} $ST_2.{\sf top}().1st = i$}{

@@

			$fp := fp \cdot \delta + x[i] \! \mod \! P$ \hspace{1cm} // $x[n]$ is the virtual character
		} \label{alg:1:d}

		sort the tuples in $ST_1'$, $ST_2'$ and $ST_3'$ by the 2nd component, respectively; \label{alg:1:e}

@@

				\Return false \hspace{1cm} // condition 2 or 3 is violated
			}	
		} \label{alg:1:g}

@@

The last point is how to obtain $\delta^{lcp[i]}$ quickly when computing $\hat{fp_1}$ and $\hat{fp_2}$ in lines~\ref{alg:1:j} and~\ref{alg:1:k}. One way is to keep a lookup table in internal memory to store all $\delta^{lcp[i]}$. This can answer the question in constant time, but it is space-consuming and impractical to be used in external memory. Notice that the LCP of any two suffixes is shorter than $n$, we can return the answer in $\mathcal{O}(\lceil \log_2{n} \rceil)$ time using $\mathcal{O}(\lceil \log_2{n} \rceil)$ internal memory. Let $e$ be an integer from $[0, n)$, its binary form is $k_{\lceil \log_2{n} \rceil}...k_1k_0$. We have $\delta^e = \Pi_{i = 0}^{\lceil \log_2{n} \rceil}{\delta}^{k_i \cdot 2^i}$, which can be computed with $\{{\delta}^{1}, {\delta}^{2}, \dots, {\delta}^{2^{\lceil \log_2{n} \rceil}} \}$ already known.

@@

Algorithm~\ref{alg:1} performs multiple scans and sorts on the arrays of $\mathcal{O}(n)$ fixed-size tuples in disks. Given RAM size $M$, disk size $D$ and block size $B$, all are in words, the time and I/O complexities for each scan are $\mathcal{O}(n)$ and $\mathcal{O}(n / B)$, respectively, while those for each sort are $\mathcal{O}(n\log_{M/ B}(n / B))$ and $\mathcal{O}((n / B)\log_{M / B}(n / B))$, respectively~\cite{Arge2013}. Algorithm~\ref{alg:1} reaches its peak disk use when sorting the tuples in lines~\ref{alg:1:b} and~\ref{alg:1:e}. Suppose the input string and the suffix/LCP array are encoded as $\alpha$- and $\beta$-byte integers, respectively, and each fingerprint is a $\gamma$-byte integer, it takes $(2 \cdot \beta + \gamma)$ space for sorting the tuples of $ST_1/ST_1'$ and $(\alpha + 2 \cdot \beta + \gamma)$ for sorting the tuples of $ST_2/ST_2'$ and $ST_3/ST_3'$. For saving space, our program implementing the algorithm tackles $ST_1/ST_1'$, $ST_2/ST_2'$ and $ST_3/ST_3'$ separately and performs a single scan over $x$ for each of them to obtain the fingerprints, using less space but more time. The experiment in Section~\ref{sec:experiment} indicates that the disk use is 40 times of $x$.

@@

We further to give another checking method using the induced sorting principle \cite{Nong11,Dementiev08}, which requires much less space than Method A. For presentation convenience, we introduce some symbols and notations as below.

Character and suffix classification. All the characters in $x$ are classified into three types, namely L-, S- and S*-type. In detail, $x[i]$ is L-type if (1) $i = n - 1$ or (2) $x[i] > x[i + 1]$ or (3) $x[i] = x[i + 1]$ and $x[i + 1]$ is L-type; otherwise, $x[i]$ is S-type. Further, if $x[i]$ and $x[i + 1]$ are separately L-type and S-type, then $x[i + 1]$ is also S*-type. Moreover, a suffix is L-, S- or S*-type if its heading character is L-, S-, or S*-type, respectively.

Suffix and LCP buckets. All the suffixes in $sa$ are partitioned into multiple buckets and those of a common heading character are grouped into a single bucket that occupies a contiguous interval in $sa$. Each bucket can be further divided into two sub-buckets, where the left and the right parts  contain L- and S-type suffixes only, respectively. For short, we use ${\sf sa\_bkt}(c)$ to denote the bucket storing the suffixes starting with character $c$ and ${\sf sa\_bkt_L}(c)/{\sf sa\_bkt_S}(c)$ to denote its left/right sub-bucket. Accordingly, $lcp$ can be also split into multiple buckets, where ${\sf lcp\_bkt}(c)/{\sf lcp\_bkt_L}(c)/{\sf lcp\_bkt_S}(c)$ stores the LCP-values of suffixes in ${\sf sa\_bkt}(c)/{\sf sa\_bkt_L}(c)/{\sf sa\_bkt_S}(c)$, respectively.

Suffix and LCP arrays for S*-type suffixes. Given that the number of S*-type suffixes is $n_1$, $sa^*[0, n_1)$ stores all the S*-type suffixes arranged in lexical order, while $lcp^*[0] = 0$ and $lcp^*[i]$ records the LCP-value of ${\sf suf}(sa^*[i])$ and ${\sf suf}(sa^*[i - 1])$ for $i \in [1, n_1)$.

Type Array. The array $t$ records in $t[i]$ the type information of $x[i]$, where $t[i] = 1$ or $0$ if $x[i]$ is S- or L-type, respectively.

@@

The induced sorting principle has been extensively used to design efficient algorithms for constructing the suffix and LCP arrays in internal or external memory \cite{Nong11,Nong14, Nong15, Bingmann12,Fischer11}. Such a construction algorithm mainly consists of a reduction phase for computing $sa^*$ and $lcp^*$, followed by an induction phase for inducing $sa$ and $lcp$ from $sa^*$ and $lcp^*$~\footnote{An overview of the induction phase is given in Appendix~\ref{sec:appendix}.}. Given that $sa^*$ and $lcp^*$ are already known, we can induce the final suffix and LCP arrays from them. This suggests a checking method based on Lemma~\ref{lemma:2}.

@@

\begin{enumerate}[(1)]
	\item
	Both $sa^*$ and $lcp^*$ are correct.
	\item
	$sa = sa'$ and $lcp = lcp'$, where $sa'$ and $lcp'$ are induced from $sa^*$ and $lcp^*$ by the IS method.
\end{enumerate}
\end{Lemma}

We have Corollary~\ref{corollary:2} for probabilistically checking the conditions of Lemma \ref{lemma:2}.

@@

	sort the tuples in $ST_1$ by the 1st component; \label{alg:2:b}

@@

	sort the tuples in $ST_2$ by the 2nd component; \label{alg:2:g}

@@

Assume that the alphabet $\Sigma$ is of size $\mathcal{O}(1)$, we can check $sa$ and $lcp$ without storing the induced suffix and LCP arrays in Algorithm~\ref{alg:2}. The idea is to compare the induced suffix/LCP items with their corresponding items in $sa/lcp$ during the inducing process. Specifically, when a suffix/LCP item $v_1$ is induced into a bucket, we check if it is equal to the corresponding item $v_2$ in $sa/lcp$. If $v_1 = v_2$, then $v_2$ is correct and we further use this value to induce the remaining suffix/LCP items. The key point is to quickly retrieve the items of $sa/lcp$ in external memory. This can be done by conducting sequential I/O operations if we provide a read pointer together with a buffer for each suffix/LCP sub-bucket. We describe below more details of the modified inducing process, where $lp_1/lp_2$ and $sp_1/sp_2$ indicate the next items to be visited in the L-type and S-type sub-buckets, respectively.
